---
title: "Data Mining- Project Working Code"
author: "Rebecca Kotula"
date: "`r Sys.Date()`"
output: pdf_document
---

# Libraries
```{r, include=FALSE}
library(ggplot2)
library(MASS) 
library(Rtsne)
library(caret)
library(GGally)
library(reticulate)
library(plyr)
library(tidyverse)
library(Rfast)
use_virtualenv("/Users/rebeccakotula/Documents/jhu/datamining/dm")
np <- import("numpy")
coresets <- import("coresets")

seed <- 1234
```


# Helper Functions
```{r}
### K-means helper functions 
predict.kmeans <- function(km.object, newdata){
    centers <- km.object$centers
    n.centers <- nrow(centers)
    dist.mat <- as.matrix(dist(rbind(centers, newdata)))
    dist.mat <- dist.mat[-seq(n.centers), seq(n.centers)]
    max.col(-dist.mat)
}

euclidean.dist <- function(a, b){
    sqrt(sum((a - b)^2))
}

kmeans.tot.withinss <- function(km.object, data){
    centers <- km.object$centers
    dist.mat <- data.frame(matrix(NA, nrow = nrow(data), ncol = nrow(centers)))
    for (i in 1:nrow(centers)){
        dist.mat[,i] <- apply(data, 1, function(x) euclidean.dist(x,centers[i,])) 
    }
    dists <- apply(dist.mat, 1, function(x) min(x))
    sum(dists^2)
}

relabel.clusters <- function(km.object, data.x, data.y, preds){
    centers <- km.object$centers
    n.centers <- nrow(centers)
    dist.mat <- as.matrix(dist(rbind(centers, data.x)))
    dist.mat <- dist.mat[-seq(n.centers), seq(n.centers)]
    min.idxs <- as.integer(colMins(dist.mat)) #rownames(dist.mat)[colMins(dist.mat)]
    clusts <- seq(1, n.centers, by=1)
    labels <- data.y[min.idxs]
    
    preds <- as.data.frame(preds)
    for (i in 1:n.centers){
        preds <- data.frame(preds)
        preds[preds$preds==i,] <- paste0(labels[i])
    }
    return(preds)
}
```

# Iris Dataset
## Data Preparation
```{r}
data(iris)
train.size <- floor(0.75 * nrow(iris))
set.seed(seed)
train.ind <- sample(seq_len(nrow(iris)), size = train.size)

train.iris <- iris[train.ind, ]
test.iris <- iris[-train.ind, ]
```


```{r}
g <- ggplot(train.iris, aes(x=Petal.Length, y = Petal.Width, col=Species)) + 
    geom_point() + 
    theme_minimal() +
    ggtitle("Iris Training Dataset with True Class Labels")
g
```

Results of K-means clustering with 3 clusters on Iris dataset (petal length and petal width features only)


## K-Means Clustering
```{r}
set.seed(seed)
km <- kmeans(train.iris[,c("Petal.Length", "Petal.Width")], 3, nstart = 25)

preds <- km$cluster
preds <- relabel.clusters(km, 
                          iris[,c("Petal.Length", "Petal.Width")], 
                          iris[,"Species"], 
                          preds)

k <- ggplot(cbind(train.iris, preds), aes(x=Petal.Length, y = Petal.Width, col=preds)) + 
    geom_point()+ 
    theme_minimal() + 
    ggtitle("Iris Training Dataset with K-Means Predicted Class Labels") +
    labs(color='Predicted Species') 
k
```

```{r iris-train-accuracy}
# Training Accuracy
acc <- sum(preds==train.iris$Species)/nrow(train.iris)
acc
```

## Coreset Construction
```{r}
py_set_seed(seed, disable_hash_randomization = TRUE)
X <- as.matrix(train.iris[,c("Petal.Length", "Petal.Width")])
cs <- coresets$KMeansCoreset(X)
coreset_gen <- cs$generate_coreset(as.integer(nrow(X)*0.1))

train.iris.coreset <- data.frame(coreset_gen[[1]])
train.iris.coreset["weights"] <- coreset_gen[[2]]
colnames(train.iris.coreset) <- c("Petal.Length", "Petal.Width","weights")

print(nrow(train.iris.coreset))
```

```{r iris-coreset-plot}
c <- ggplot() + geom_point(data=train.iris, aes(x=Petal.Length, y = Petal.Width)) + 
    geom_point(data=train.iris.coreset, aes(x=Petal.Length,y=Petal.Width, size=weights), alpha=0.6,col='red')+ 
    theme_minimal() + 
    ggtitle("Iris Training Dataset with Coreset in Red")
c
```

## K-Means on Coreset
```{r iris-coreset-kmeans}
set.seed(seed)
km.c <- kmeans(train.iris.coreset[,1:2], 3, nstart = 25)
preds.c <- km.c$cluster
preds.c <- relabel.clusters(km.c, 
                          iris[,c("Petal.Length", "Petal.Width")], 
                          iris[,"Species"], 
                          preds.c)

train.iris.coreset["Species.Pred"] <- preds.c

k.c <- ggplot(train.iris.coreset, aes(col=Species.Pred)) +
    geom_point(aes(x=Petal.Length,y=Petal.Width))+ 
    theme_minimal() + 
    ggtitle("Coreset with K-means Predicted Species")
k.c
```

```{r kmeans-iris-centers}
c <- ggplot() + geom_point(data=iris, aes(x=Petal.Length, y = Petal.Width, col=Species)) + 
    geom_point(aes(x=km$centers[,1],y=km$centers[,2]), col='blue', shape=15, size=3)+
    geom_point(aes(x=km.c$centers[,1],y=km.c$centers[,2]), col='red', shape=15, size=3)+ 
    theme_minimal() +
    ggtitle("K-means Centers on Full Iris Dataset") 
c
```


## Results

### Accuracy and Confusion Matrix 

```{r iris-test-accuracy}
set.seed(seed)
test.preds <- predict.kmeans(km, test.iris[,c("Petal.Length","Petal.Width")])
test.preds <- relabel.clusters(km, 
                          iris[,c("Petal.Length", "Petal.Width")], 
                          iris[,"Species"], 
                          test.preds)

table(test.preds$preds, test.iris$Species)
```

```{r}
acc.test <- sum(test.preds$preds==test.iris$Species)/nrow(test.iris)
acc.test
```

```{r iris-coreset-test-accuracy}
set.seed(seed)
test.preds.c <- predict.kmeans(km.c, test.iris[,c("Petal.Length","Petal.Width")])
test.preds.c <- relabel.clusters(km.c, 
                          iris[,c("Petal.Length", "Petal.Width")], 
                          iris[,"Species"], 
                          test.preds.c)

table(test.preds.c$preds, test.iris$Species)
```
```{r}
acc.test.c <- sum(test.preds.c$preds==test.iris$Species)/nrow(test.iris)
acc.test.c
```

### tot.withinss
```{r iris-withinss}
kmeans.tot.withinss(km,iris[,3:4])
```

```{r iris-coreset-withinss}
kmeans.tot.withinss(km.c,iris[,3:4])
```


## Coreset Size Sweep

```{r}
sizes <- c(0.01,0.04, 0.05, 0.1, 0.25)
py_set_seed(seed, disable_hash_randomization = TRUE)
coreset.04 <- data.frame((cs$generate_coreset(as.integer(nrow(X)*0.04)))[[1]])
coreset.05 <- data.frame((cs$generate_coreset(as.integer(nrow(X)*0.05)))[[1]])
coreset.1 <- data.frame((cs$generate_coreset(as.integer(nrow(X)*0.1)))[[1]])
coreset.25 <- data.frame((cs$generate_coreset(as.integer(nrow(X)*0.25)))[[1]])

colnames(coreset.04) <- c("Petal.Length", "Petal.Width")
colnames(coreset.05) <- c("Petal.Length", "Petal.Width")
colnames(coreset.1) <- c("Petal.Length", "Petal.Width")
colnames(coreset.25) <- c("Petal.Length", "Petal.Width")

set.seed(seed)
km.c.04<- kmeans(coreset.04, 3, nstart = 25)
km.c.05<- kmeans(coreset.05, 3, nstart = 25)
km.c.1<- kmeans(coreset.1, 3, nstart = 25)
km.c.25<- kmeans(coreset.25, 3, nstart = 25)

test.preds.04 <- predict.kmeans(km.c.04, test.iris[,c("Petal.Length","Petal.Width")])
test.preds.05 <- predict.kmeans(km.c.05, test.iris[,c("Petal.Length","Petal.Width")])
test.preds.1 <- predict.kmeans(km.c.1, test.iris[,c("Petal.Length","Petal.Width")])
test.preds.25 <- predict.kmeans(km.c.25, test.iris[,c("Petal.Length","Petal.Width")])

test.preds.04 <- relabel.clusters(km.c.04,
                               iris[,c("Petal.Length", "Petal.Width")],
                               iris[,"Species"],
                               test.preds.04)
test.preds.05 <- relabel.clusters(km.c.05,
                               iris[,c("Petal.Length", "Petal.Width")],
                               iris[,"Species"],
                               test.preds.05)
test.preds.1 <- relabel.clusters(km.c.1,
                               iris[,c("Petal.Length", "Petal.Width")],
                               iris[,"Species"],
                               test.preds.1)
test.preds.25 <- relabel.clusters(km.c.25,
                               iris[,c("Petal.Length", "Petal.Width")],
                               iris[,"Species"],
                               test.preds.25)

test.acc.04<- sum(test.preds.04$preds==test.iris$Species)/nrow(test.iris)
test.acc.05<- sum(test.preds.05$preds==test.iris$Species)/nrow(test.iris)
test.acc.1<- sum(test.preds.1$preds==test.iris$Species)/nrow(test.iris)
test.acc.25<- sum(test.preds.25$preds==test.iris$Species)/nrow(test.iris)

test.acc <- c("--", test.acc.04, test.acc.05, test.acc.1, test.acc.25)

totwss.04 <- kmeans.tot.withinss(km.c.04,train.iris[,3:4])
totwss.05 <- kmeans.tot.withinss(km.c.05,train.iris[,3:4])
totwss.1 <- kmeans.tot.withinss(km.c.1,train.iris[,3:4])
totwss.25 <- kmeans.tot.withinss(km.c.25,train.iris[,3:4])

totwss <- c("--", totwss.04, totwss.05, totwss.1, totwss.25)

iris.sweep <- data.frame(coreset.size = sizes, test.accuracy = test.acc, tot.w.ss = totwss)
```


# Simulated Data

## Data Preparation 

```{r}
# set seed and create data vectors 
set.seed(seed) 
sample.size <- 4000                                        
class1.mu <- c(10, 5, 7, 9, 20)   
class2.mu <- c(9, 3, 4, 5, 15)   
class3.mu <- c(3, 2, 1, 4, 9)   
class4.mu <- c(5, 5, 5, 5, 5)   

n <- 5  
A1 <- matrix(runif(n^2)*2-1, ncol=n) 
sigma1 <- t(A1) %*% A1

A2 <- matrix(runif(n^2)*4-2, ncol=n) 
sigma2 <- t(A2) %*% A2

A3 <- matrix(runif(n^2)*3-2, ncol=n) 
sigma3 <- t(A3) %*% A3

A4 <- matrix(runif(n^2)*2-2, ncol=n) 
sigma4 <- t(A4) %*% A4
  
# create multivariate normal distribution 
class1.dat <- data.frame(mvrnorm(n = sample.size, 
                               mu = class1.mu,  
                               Sigma = sigma1))
class2.dat <- data.frame(mvrnorm(n = sample.size, 
                               mu = class2.mu,  
                               Sigma = sigma2)) 
class3.dat <- data.frame(mvrnorm(n = sample.size, 
                               mu = class3.mu,  
                               Sigma = sigma3))
class4.dat <- data.frame(mvrnorm(n = sample.size, 
                               mu = class4.mu,  
                               Sigma = sigma4))

class1.dat$class <- "a"
class2.dat$class <- "b"
class3.dat$class <- "c"
class4.dat$class <- "d"

sim.data <- rbind(class1.dat,rbind(class2.dat,rbind(class3.dat, class4.dat)))

sim.mat <- as.matrix(sim.data[,-6]) 

sim.sd <- apply(sim.mat, 1, sd)
sim.dat <- sim.mat + rnorm(length(sim.mat), mean = 0, sd = sim.sd * 0.05)
sim.data[,1:5] <- sim.dat
```


```{r}
train.size.sim <- floor(0.75 * nrow(sim.data))
set.seed(seed)
train.ind.sim <- sample(seq_len(nrow(sim.data)), size = train.size.sim)

train.sim <- sim.data[train.ind.sim, ]
test.sim <- sim.data[-train.ind.sim, ]
print(nrow(train.sim))
print(nrow(test.sim))
```



```{r}
set.seed(seed)
train.sim.mat <- as.matrix(train.sim[,-6])
tsne.sim.out <- Rtsne(train.sim.mat, check_duplicates = FALSE)
```

```{r}
tsne.sim.plot <- data.frame(x = tsne.sim.out$Y[,1],
                        y = tsne.sim.out$Y[,2])
g.sim.tsne <- ggplot(tsne.sim.plot) +
    geom_point(aes(x=x, y=y, col = as.factor(train.sim[,6]))) +
    theme_minimal() +
    labs(color='class') +
    ggtitle("t-SNE Representation of Simulated Training Dataset")
g.sim.tsne
```

## K-Means Clustering
```{r}
set.seed(seed)
km.sim<- kmeans(train.sim[,-6], 4, nstart = 25)

preds.sim <- km.sim$cluster
preds.sim <- relabel.clusters(km.sim, train.sim[,-6], train.sim[,6], preds.sim)

sim.tsne.pred <- ggplot(cbind(tsne.sim.plot, preds.sim)) +
    geom_point(aes(x=x, y=y, col = preds)) +
    theme_minimal() +
    labs(color='class') +
    ggtitle("t-SNE Representation of Simulated Training Dataset K-Means Predictions")
sim.tsne.pred
```

```{r sim-training-accuracy}
acc.sim <- sum(preds.sim$preds==train.sim[,6])/nrow(train.sim)
acc.sim
```


```{r}
table(as.factor(preds.sim$preds), as.factor(train.sim[,6]))
```
## Test Accuracy 
```{r}
set.seed(seed)
test.sim.preds <- predict.kmeans(km.sim, test.sim[,-6])
test.sim.preds <- relabel.clusters(km.sim, 
                                   train.sim[,-6], 
                                   train.sim[,6], 
                                   test.sim.preds)

table(test.sim.preds$preds, test.sim[,6])
```

```{r}
acc.test.sim <- sum(test.sim.preds$preds==test.sim[,6])/nrow(test.sim)
acc.test.sim
```

## Coreset Construction

```{r sim-coreset}
# Generate Coreset
py_set_seed(seed, disable_hash_randomization = TRUE)
X.sim <- as.matrix(train.sim[,-6])
cs.sim <- coresets$KMeansCoreset(X.sim, n_clusters = as.integer(4), random_state = as.integer(1234))
coreset_gen.sim <- cs.sim$generate_coreset(as.integer(length(X.sim)*0.01))

train.sim.coreset <- data.frame(coreset_gen.sim[[1]])
print(nrow(train.sim.coreset))

# Which indices from original dataset are in the coreset: 
train.sim.list <- paste(train.sim[,1], train.sim[,2], train.sim[,3], train.sim[,4], train.sim[,5])
train.sim.coreset.list <- paste(train.sim.coreset[,1], train.sim.coreset[,2], train.sim.coreset[,3], train.sim.coreset[,4], train.sim.coreset[,5])
coreset.ind <- which(train.sim.list %in% train.sim.coreset.list)
```

```{r}
c.sim <- ggplot() + 
    geom_point(data=tsne.sim.plot, aes(x=x, y=y), alpha=0.6) + 
    geom_point(data=coreset.plot, aes(x=x,y=y),  col='red')+ 
    theme_minimal() + 
    ggtitle("Simulated Training Dataset with 1% Coreset in Red")
c.sim
```


## K-Means on Coreset

```{r}
set.seed(seed)
km.c.sim <- kmeans(train.sim.coreset, 4, nstart = 25)
preds.c.sim <- km.c.sim$cluster
preds.c.sim <- relabel.clusters(km.c.sim, 
                                train.sim[,-6], 
                                train.sim[,6], 
                                preds.c.sim)

k.c.sim <- ggplot(cbind(tsne.sim.plot[coreset.ind,], preds.c.sim)) +
    geom_point(aes(x=x,y=y, col=preds))+ 
    theme_minimal() + 
    ggtitle("Coreset with K-means Predicted Class")
k.c.sim
```

```{r}
c.sim <- ggplot() + #geom_point(data=tsne.sim.plot, aes(x=x, y = y, col=as.factor(train.sim[,6]))) + 
    geom_point(aes(x=km.sim$centers[,1],y=km.sim$centers[,2]), col='blue', shape=15, size=3)+
    geom_point(aes(x=km.c.sim$centers[,1],y=km.c.sim$centers[,2]), col='red', shape=15, size=3)+ 
    theme_minimal() +
    ggtitle("K-means Centers on Simulated Training Dataset") 
c.sim
```


## Coreset Size Sweep

```{r simulated-coreset-size-sweep}
py_set_seed(seed, disable_hash_randomization = TRUE)
coreset.01.sim <- data.frame((cs.sim$generate_coreset(as.integer(nrow(X.sim)*0.01)))[[1]])
coreset.04.sim <- data.frame((cs.sim$generate_coreset(as.integer(nrow(X.sim)*0.04)))[[1]])
coreset.05.sim <- data.frame((cs.sim$generate_coreset(as.integer(nrow(X.sim)*0.05)))[[1]])
coreset.1.sim <- data.frame((cs.sim$generate_coreset(as.integer(nrow(X.sim)*0.1)))[[1]])
coreset.25.sim <- data.frame((cs.sim$generate_coreset(as.integer(nrow(X.sim)*0.25)))[[1]])

set.seed(seed)
km.c.01.sim <- kmeans(coreset.01.sim, 4, nstart = 25)
km.c.04.sim <- kmeans(coreset.04.sim, 4, nstart = 25)
km.c.05.sim <- kmeans(coreset.05.sim, 4, nstart = 25)
km.c.1.sim <- kmeans(coreset.1.sim, 4, nstart = 25)
km.c.25.sim <- kmeans(coreset.25.sim, 4, nstart = 25)

test.preds.01.sim <- predict.kmeans(km.c.01.sim, test.sim[,-6])
test.preds.04.sim <- predict.kmeans(km.c.04.sim, test.sim[,-6])
test.preds.05.sim <- predict.kmeans(km.c.05.sim, test.sim[,-6])
test.preds.1.sim <- predict.kmeans(km.c.1.sim, test.sim[,-6])
test.preds.25.sim <- predict.kmeans(km.c.25.sim, test.sim[,-6])

test.preds.01.sim <- relabel.clusters(km.c.01.sim,
                                test.sim[,-6],
                                test.sim[,6],
                               test.preds.01.sim)
test.preds.04.sim <- relabel.clusters(km.c.04.sim,
                                test.sim[,-6],
                                test.sim[,6],
                               test.preds.04.sim)
test.preds.05.sim <- relabel.clusters(km.c.05.sim,
                                test.sim[,-6],
                                test.sim[,6],
                               test.preds.05.sim)
test.preds.1.sim <- relabel.clusters(km.c.1.sim,
                                test.sim[,-6],
                                test.sim[,6],
                               test.preds.1.sim)
test.preds.25.sim <- relabel.clusters(km.c.25.sim,
                                test.sim[,-6],
                                test.sim[,6],
                               test.preds.25.sim)

test.acc.01.sim <- sum(test.preds.01.sim$preds==test.sim[,6])/nrow(test.sim)
test.acc.04.sim <- sum(test.preds.04.sim$preds==test.sim[,6])/nrow(test.sim)
test.acc.05.sim <- sum(test.preds.05.sim$preds==test.sim[,6])/nrow(test.sim)
test.acc.1.sim <- sum(test.preds.1.sim$preds==test.sim[,6])/nrow(test.sim)
test.acc.25.sim <- sum(test.preds.25.sim$preds==test.sim[,6])/nrow(test.sim)

test.acc <- c(test.acc.01.sim, 
              test.acc.04.sim, 
              test.acc.05.sim, 
              test.acc.1.sim, 
              test.acc.25.sim)

totwss.01.sim <- kmeans.tot.withinss(km.c.01.sim,sim.data[,-6])
totwss.04.sim <- kmeans.tot.withinss(km.c.04.sim,sim.data[,-6])
totwss.05.sim <- kmeans.tot.withinss(km.c.05.sim,sim.data[,-6])
totwss.1.sim <- kmeans.tot.withinss(km.c.1.sim,sim.data[,-6])
totwss.25.sim <- kmeans.tot.withinss(km.c.25.sim,sim.data[,-6])

totwss <- c(totwss.01.sim, totwss.04.sim, totwss.05.sim, totwss.1.sim, totwss.25.sim)

sim.sweep <- data.frame(coreset.size = sizes, test.accuracy = test.acc, tot.w.ss = totwss)
```